SDM\_the\_search\_for\_powers\_of\_2\_everywhere

### 2018-04-27 Chapters/Chapter02.tex
sdm ready?

% TODO GARY KLEIN !!!
GARY KLEIN!!!!

Suddenly, you feel the danger.  You tighten your grip on the child's hand and step up your walking.  In these large cities, walking on a dark alley was never recommended, but somehow that's what you did, and now everthing you think of is to get out to safety.  You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exactly is causing your fear. One just feels it. We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind you; a general lack of activity; that very feeling that something isn't right... without knowing precisely what isn't right, or even what `right' means... etc. --- recalling past experiences from memory and thus generating the feeling. Our memory is able to make a parallel between previous experiences and the clues. Although one has never been in the exactly same situation, one's brain involuntarily makes an analogy and recognizes the possibility of danger. The whole process happens so fast it feels like it only took a single unit of psychological time.

Let us compare this to what a child might do.

@@

Sparse Distributed Memory reads like a \emph{desiderata} of a theory for human memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
\begin{enumerate}
    \item Why are most concepts orthogonal, unrelated to each other?
    \item Why is there Miller's magic number, i.e., we can't hold too many things in mind at once?
    \item Why do we at times instantly recall an experience; other times we can't recall anything at all; and still other times we get into this strange tip-of-the-tongue situation... it is clearly `there'... but where?  (and `what' is `there'?)
    \item How does this recall process work?  What is remembering?
    \item Why do neurons die and we still remember most everything?
    \item What do neurons actually do?
\end{enumerate}

@@

\section{The wasted effort of duplicated, AD-HOC, work}

@@

    \begin{quote}
    The most exciting phrase to hear in science, the one that heralds new discoveries, is not `Eureka!' (I found it!) but ``That's funny...''
    — Isaac Asimov
    \end{quote}

    \bigskip
\begin{quote}
Alexander Fleming, while explaining the staphylococcal work to Pryce, to demonstrate a point, picked up one of the top layer of the discarded plates when he noticed the lysis of the staphylococci (his cry of Eureka was ‘That’s funny!’) and then and there made a subculture of the mould.
\end{quote}

@@

\begin{enumerate}
  \item Calculate critical distance.ipynb
  \item Calculate radius for any SDM.ipynb
  \item Classification Test 1.ipynb
  \item Classification Test 2.ipynb
  \item Classification Test 3-Copy1.ipynb
  \item Classification Test 3.ipynb
  \item Critical Distance - New.ipynb
  \item Critical Distance.ipynb
  \item Distance between bitstrings.ipynb
  \item Distances of activated hard-locations.ipynb
  \item Generates a new SDM.ipynb
  \item index.rst
  \item Intersection of activated hard-locations.ipynb
  \item Kanerva-Table-7 - Experiment.ipynb
  \item Kanerva-Table-7.3 - Generic Read.ipynb
  \item Kanerva-Table-7.3 - Multiple dimensions.ipynb
  \item Kanerva-Table-7.3 - Weighted table write.ipynb
  \item Kanerva-Table-7.3.ipynb
  \item Kanerva's Figure 1.2.ipynb
  \item Noise filter 2.ipynb
  \item Noise filter A-Z 0-1 Generator.ipynb
  \item Noise filter with labels.ipynb
  \item Noise filter.ipynb
  \item Number of activated hard-locations.ipynb
  \item Performance test.ipynb
  \item Reseting hard-locations.ipynb
  \item Sequences (Kanerva Ch 8).ipynb
  \item Sequences in one SDM (Kanerva Ch 8).ipynb
  \item table74.pickle
  \item TicTacToe - Old (TO BE DELETED).ipynb
  \item TicTacToe.ipynb
  \item tmp-TO-BE-DELETED-2.ipynb
  \item tmp-TO-BE-DELETED.ipynb
  \item Untitled.ipynb
  \item Untitled1.ipynb
  \item Weighted operations using information.ipynb
\end{enumerate}

>>>>

% TODO GARY KLEIN !!!

Suddenly, you feel the danger.  You tighten your grip on the child's hand and step up your walking.  In these large cities, walking on a dark alley was never recommended, but somehow that's what you have done --- and now everything you can think of is to get back to safety.  You are afraid of something, but you cannot explain --- in rigorous detail, at least --- what exactly is causing your fear. One just feels it.

We may interpret this situation as clues of the present --- a dark alley; a giant metropolitan area; people going on about their lives mostly indifferent from each other; constrained routes ahead and behind; a general lack of activity; that very feeling that something isn't right... without knowing precisely what isn't right, or even what `right' means --- recalling past experiences from memory and thus generating the feeling. Our memory is able to make a parallel between previous experiences and the clues. Although one has never been in the exactly same situation, one's brain involuntarily, through analogy with previous experiences, recognizes the possibility of danger. The whole process happens so fast it feels like it only took a single indivisible unit of psychological time.  A single cycle, to use the computing word.

As much as we can rationalize our ability to understand the why of a feeling like this, it seems unlikely that we will be able to nail all reasons down into rules.  Let us compare this to what a child might do.  While an adult may respond by discreetly stepping up the walking speed, a child might open her eyes and be surprised and ask, indiscretely, questions about the walking speed... Because children do not have the accumulated experience of adults in dealing with large metropolian areas, their behavior differs markedly from ours.

@@

Sparse Distributed Memory reads like a \emph{desiderata} of a theory for human memory.  To understand the breadth of topics that SDM encompasses, consider the following questions:
\begin{enumerate}
    \item Why are most concepts orthogonal, unrelated to each other?
    \item Why is there Miller's magic number, i.e., we can't hold too many things in mind at once?
    \item Why do we at times instantly recall an experience; other times we can't recall anything at all; and still other times we get into this strange tip-of-the-tongue situation... it is clearly `there'... but where?  (and `what' is `there'?\footnote{Psychologists have documented interesting properties about this state:  it happens, for instance, around every 1-week cycle; and it happens mostly over proper names, subjects have access to the first letter and to the number of syllables better than chance, etc. Also interesting is their clever way of triggerring the state: through definitions of rarely-used words.  For some psychological studies about the tip-of-tongue, see \citet{meyer_tip---tongue_1992, brown_review_1991, brown_tip_1966}.}.)
    \item How does this recall process work?  What is remembering?
    \item Why do neurons die and we still remember most everything?
    \item What do neurons actually do?  What is their primary function?
\end{enumerate}

@@

\section{The wasted effort of duplicated, ad hoc, work}

@@

    \begin{quote}
    The most exciting phrase to hear in science, the one that heralds new discoveries, is not `Eureka!' (I found it!) but ``That's funny...'' \\
    — Isaac Asimov
    \end{quote}

@@

\begin{enumerate}
  \item Calculate critical distance.ipynb
  \item Calculate radius for any SDM.ipynb
  \item Classification Test 1.ipynb
  \item Classification Test 2.ipynb
  \item Classification Test 3-Copy1.ipynb
  \item Classification Test 3.ipynb
  \item Critical Distance - New.ipynb
  \item Critical Distance.ipynb
  \item Distance between bitstrings.ipynb
  \item Distances of activated hard-locations.ipynb
  \item Generates a new SDM.ipynb
  \item Intersection of activated hard-locations.ipynb
  \item Kanerva-Table-7 - Experiment.ipynb
  \item Kanerva-Table-7.3 - Generic Read.ipynb
  \item Kanerva-Table-7.3 - Multiple dimensions.ipynb
  \item Kanerva-Table-7.3 - Weighted table write.ipynb
  \item Kanerva-Table-7.3.ipynb
  \item Kanerva's Figure 1.2.ipynb
  \item Noise filter 2.ipynb
  \item Noise filter A-Z 0-1 Generator.ipynb
  \item Noise filter with labels.ipynb
  \item Noise filter.ipynb
  \item Number of activated hard-locations.ipynb
  \item Performance test.ipynb
  \item Reseting hard-locations.ipynb
  \item Sequences (Kanerva Ch 8).ipynb
  \item Sequences in one SDM (Kanerva Ch 8).ipynb
  \item TicTacToe.ipynb
  \item Weighted operations using information.ipynb
\end{enumerate}
